---
title: "Evil Eye Analysis Project for Github"
author: "Madeleine Zoeller"
date: "`r Sys.Date()`"
output: html_document
---


##Setting up

Restart
```{r}
rm(list= ls())
```

Loading packages
```{r}
library(umx)
library(psych)
library(matrixStats)
library(effsize)
library(psycho)
library(finalfit)
library(xtable)
library(stargazer)
library(MASS)
library(MuMIn)
library(estimatr)
library(corpcor)
library(GGally)
library(AID)
library(faraway)
library(plyr)
library(mvnmle)
library(BaylorEdPsych)
library(mice)
library(lme4)
library(readxl)
library(gvlma)
library(dplyr)
library(tidyverse)
library(moderndive)
library(boxcoxmix)
library(car)
library(semPlot)
library(lavaan)
library(caret)
library(car)
library(faraway)
library(alr3)
library(ggcorrplot)


```
Reading in data
```{r}
ee_data <- read.csv("p5_english_num_EE_data_5_7_20_csv.csv", header = T, na.strings=c(""," ","<","NA"))
```
## Data Cleaning and processing

Number of people from each country who were paid each amount
```{r}
#Need number of participants by country (before bot check 483)

country <- select(ee_data, language)

fr <-  ee_data %>% filter(language == "French") %>% nrow()

fr

#195 french

sp <-  ee_data %>% filter(language == "Spanish") %>% nrow()
sp

#261 spanish

por <-  ee_data %>% filter(language == "Portuguese") %>% nrow()

por
#18 portuguese

pol <-  ee_data %>% filter(language == "Polish") %>% nrow()

pol
#9 polish
```

Removing subjects who failed the bot check
```{r}
nrow(ee_data)
view(ee_data)
#removing rows that failed the bot check question
ee_data_bc <- ee_data %>%
  filter( botcheck == "None" )
View(ee_data_bc)
nrow(ee_data_bc)
ncol(ee_data_bc)
#50 people failed the bot check, 433 remain


```
Removing tedious variables and changing certain NAs
```{r}
View(ee_data_bc)
ee_data_bc <- ee_data_bc[, -c(1,2,37:49)]

#changing NAs in children to 0 
ee_data_bc$children[is.na(ee_data_bc$children)] <- 0
sum(is.na(ee_data_bc$children))

View(ee_data_bc)

sum(is.na(ee_data_bc)) #43 NAs left

```
Removing other NAs (Data was not missing completely at random)
```{r}
#Before removing NAs
nrow(ee_data_bc)
#433 rows
sum(is.na(ee_data_bc))
#43 NAs

#At this point, the best option is probably ommitting all NAs
ee_nona_data_bc <- na.omit(ee_data_bc)

nrow(ee_nona_data_bc)

#395 rows left 

#rename to make code less clunky

ee_final <- ee_nona_data_bc
```
Adding participant ID numbers and getting demographics for methods section
```{r}
#add participant numbers
ee_final$row_num <- seq.int(nrow(ee_final))

View(ee_final)

#gender
ee_final %>% filter(gender == "Woman") %>% nrow()

#130 women
ee_final %>% filter(gender == "Man") %>% nrow()

#263 men
ee_final %>% filter(gender == "Other") %>% nrow()
#2 Other

#average age

mean(ee_final$age)

sd(ee_final$age)

#French
ee_final %>% filter(language == "French") %>% nrow()

ee_final %>% filter(language == "Spanish") %>% nrow()

ee_final %>% filter(language == "Polish") %>% nrow()

ee_final %>% filter(language == "Portuguese") %>% nrow()
```
Putting income as a numeric 
```{r}

is.data.frame(ee_final)


ee_final$income <- as.numeric(as.character(ee_final$income))

class(ee_final$income)

is.numeric(ee_final$age)

#NAs introduced by above process, code below fixes it

any(is.na(ee_final$income))

ee_final$income

str(ee_final)

#names of columns with missing values
#https://stackoverflow.com/questions/20364450/find-names-of-columns-which-contain-missing-values

colnames(ee_final)[colSums(is.na(ee_final)) > 0]

ee_final$income[is.na(ee_final$income)] <- 0

any(is.na(ee_final$income))

```
Converting Baymul Income inequality visual into gini coefficients, adding gini_17, and making gini_error
```{r}
#did it this way because of code issues with factors down the line
ee_final <- as_tibble(ee_final)


is_tibble(ee_final)


ee_final <- ee_final %>% mutate(gini_baymul_num = ifelse(gini_baymul %in% "A", .42,ifelse(gini_baymul %in% "B", .35,ifelse(gini_baymul %in% "C", .30, ifelse(gini_baymul %in% "D", .20, ifelse(gini_baymul %in% "E", .21, 0))))))
View(ee_final)




  
ee_final <- ee_final %>% mutate(gini_17 = ifelse(language %in% "Portuguese", .338,ifelse(language %in% "Spanish", .347,ifelse(language %in% "French", .316, ifelse(language %in% "Polish", .297, 0)))))


ee_final <- mutate(ee_final, gini_error = gini_baymul_num - gini_17)


```
Adjusting reverse coded questions
```{r}
ee_final <- ee_final %>% mutate(ii1_st = ifelse(ii1_rc %in% 5, 1, ifelse(ii1_rc %in% 4, 2, ifelse(ii1_rc %in% 3, 3, ifelse(ii1_rc %in% 2, 4, ifelse(ii1_rc %in% 1, 5, NA))))))

#higher values of srd3_rc and srd4_rc should become lower values
ee_final <- ee_final %>% mutate(srd3_st = ifelse(srd3_rc %in% 5, 1, ifelse(srd3_rc %in% 4, 2, ifelse(srd3_rc %in% 3, 3, ifelse(srd3_rc %in% 2, 4, ifelse(srd3_rc %in% 1, 5, 0))))))


ee_final <- ee_final %>% mutate(srd4_st = ifelse(srd4_rc %in% 5, 1, ifelse(srd4_rc %in% 4, 2, ifelse(srd4_rc %in% 3, 3, ifelse(srd4_rc %in% 2, 4, ifelse(srd4_rc %in% 1, 5, 0))))))

```
Creating Summary Variables
```{r}
#creating summary variables 

ee_final <- ee_final %>%
  mutate(bzsg_sum = bzsg1 +bzsg2 + bzsg3,
         srd_sum = srd1 + srd2 + srd3_st + srd4_st,
         ee_sum = ee1 + ee2 + ee3 + ee4 + ee5 + ee6 + ee7 + ee8 + ee9 + ee10, ii_sum = ii2 + ii1_st,  bjw_sum = bjw_nature + bjw_god + bjw_deserve + bjw_chance + bjw_nr, theft_sum = theft1 + theft2 + theft3, cii = courts1 + police1 + courts2 + police2)



#higher values of bzsg indicate higher belief, higher values of srd indicate feeling more deprived, higher values of ee indicate greater belief in ee, higher values of ii indicate high percieved inequality in their vicinity, higher bjw means more belief in a just world
#cii= confidence in institutions (court and police)

```
Converting income to USD using PPP then creating income per household member in PPP dollars
```{r}
#PPP
#https://www.open.edu/openlearn/science-maths-technology/mathematics-and-statistics/how-compare-income-across-countries



#PPP conversion factors
#ppp_cf is ppp conversion factor

#France
#.77

#Portugal
#.59

#Poland
#1.78
#Spain
#.65

#might need to transform language responses to English 



ee_final <- mutate(ee_final, ppp_cf = ifelse(language %in% "French", .77, ifelse(language %in% "Spanish", .65, ifelse( language %in% "Portugese", .59, ifelse(language %in% "Polish", 1.78, 0)))))




ee_final <- mutate(ee_final, inc_ppp = income * ppp_cf)

ee_final <- mutate(ee_final, inc_ppp_pp = inc_ppp/household_size)



#removing NaN created 

ee_final$inc_ppp_pp[is.na(ee_final$inc_ppp_pp)] <- 0

#removing other NAs created
ee_final[is.na(ee_final)] <- 0

#removing infinity and negative infinity
ee_final$inc_ppp_pp[is.infinite(ee_final$inc_ppp_pp)] <- 0
```
## Data analysis
###Models
Model 1
```{r}
#regular model
lm1 <- lm(ee_sum ~ bzsg_sum + srd_sum + ii_sum + age + inc_ppp_pp + cii + theft_sum + language + gini_baymul + bjw_sum + children + edu, data = ee_final)

summary(lm1)



#robust model
#gini_baymul
lm1_r <- lm_robust(ee_sum ~ bzsg_sum + srd_sum + ii_sum + age + inc_ppp_pp + cii + theft_sum + language + gini_baymul + bjw_sum + children + edu + gender, data = ee_final)


#gini_error
lm2_r <- lm_robust(ee_sum ~ bzsg_sum + srd_sum + ii_sum + age + inc_ppp_pp + cii + theft_sum + language + gini_error + bjw_sum + children + edu + gender, data = ee_final)




ee_final %>% filter(language == "French") %>% nrow()
```
AIC/BIC
```{r}

library(MASS)
MASS::stepAIC(lm1, scale = 0,
        direction = c("both", "backward", "forward"),
        trace = 1, keep = NULL, steps = 1000, use.start = FALSE,
        k = 2)

#interpreting AIC values
#https://ashutoshtripathi.com/2019/06/10/what-is-stepaic-in-r/

#AIC vs BIC
#http://www.differencebetween.net/miscellaneous/difference-between-aic-and-bic/

```
New model based on stepAIC
```{r}

final_model <- lm_robust(formula = ee_sum ~ srd_sum + cii + theft_sum + language +  bjw_sum + children, data = ee_final)

summary(final_model)

r2 <- .1963

f2 <- r2/(1-r2)
f2
```
Looking at robust model more closely
```{r}
#Portugal and Poland have smaller residuals (and they are symmetric) than France and Spain

library("car")

residualPlot(lm1, groups = "language")

plot(residuals(lm1), col = ee_final$language, xlab = "language")

plot(residuals(lm1), col = ee_final$language)




residualPlots(lm1)
```





## Basic Descriptive Statistics
###Mean and SD of all numerics
```{r}

#age

mean(ee_final$age)

sd(ee_final$age)

#bzsg

mean(ee_final$bzsg_sum)

sd(ee_final$bzsg_sum)

#bjw

mean(ee_final$bjw_sum)

sd(ee_final$bjw_sum)

#cii
mean(ee_final$cii)

sd(ee_final$cii)

#children

mean(ee_final$children)

sd(ee_final$children)

#age

mean(ee_final$age)
sd(ee_final$age)

#srd

mean(ee_final$srd_sum)

sd(ee_final$srd_sum)

#local inequality

mean(ee_final$ii_sum)

sd(ee_final$ii_sum)

#experiences with theft

mean(ee_final$theft_sum)

sd(ee_final$theft_sum)

#ee sum
mean(ee_final$ee_sum)
sd(ee_final$ee_sum)
```
###Mean and SD of EE components
```{r}
is.data.frame(ee_final)
View(ee_final)
ee_items <- ee_final %>% select(ee1, ee2, ee3, ee4, ee5, ee6, ee7, ee8, ee9, ee10)


View(ee_items)
ee_i_m <- as.matrix(ee_items)
ee_means <- colMeans(ee_i_m)
ee_sds <- colSds(ee_i_m)

View(ee_means)
View(ee_sds)

View(ee_sds)
View(ee_means)


ee_sds_r <- round(ee_sds, digits = 2)
ee_means_r <- round(ee_means, digits = 2)
View(ee_sds_r)


View(ee_means_r)





library(dplyr)

##bleh
ee_items <- as.matrix(ee_items)



#merging columns 
install.packages("matrixStats")
library(matrixStats)
is.data.frame(ee_items)

view(ee_items)

colSds(ee_items)
?colSds
ee_means <- colMeans(ee_items)

mean(ee_final$ee1)
sd(ee_final$ee1)
View(ee_means)

ee_sds <- 

```
###Mean and SD of II components
```{r}
ii_items <- ee_final %>% select(ii1_rc, ii1_st, ii2, ii_sum)

View(ii_items)


#ii sum is ii1_st + ii2

mean(ii_items$ii1_st)
sd(ii_items$ii1_st)

mean(ii_items$ii2)
sd(ii_items$ii2)

##below attempts at fanciness
#get ii items

ii_items <- ee_final %>% select(ii1_rc, ii2)
View(ii_items)

is.data.frame(ii_items)
class(ii_items)

#try to make a data frame with items, mean and sd of item
 
ii_stats <- ii_items %>% summarise(mean = mean(ii_items), sd = sd(ii_items))

#try making ii_items a matrix?

ii_items_m <- as.matrix(ii_items)

class(ii_items_m)

mean(ii_items)
#did not work 

## perhaps time to switch back to other method?

##Ex mean(ii_items$ii1_rc)

mean(ii_items$ii1_rc)
```
###Mean and SD of CII components
```{r}
View(ee_final)

cii_items <- ee_final %>% select(police1, police2, courts1, courts2)

mean(cii_items$police1)
sd(cii_items$police1)

mean(cii_items$police2)
sd(cii_items$police2)

mean(cii_items$courts1)
sd(cii_items$courts1)

mean(cii_items$courts2)
sd(cii_items$courts2)
```
###Mean and SD of SRD components
```{r}
View(ee_final)

srd_items <- ee_final %>% select(srd1, srd2, srd3_st, srd4_st)

mean(srd_items$srd1)
sd(srd_items$srd1)

mean(srd_items$srd2)
sd(srd_items$srd2)

mean(srd_items$srd3_st)
sd(srd_items$srd3_st)

mean(srd_items$srd4_st)
sd(srd_items$srd4_st)
```
###Mean and SD of BZSG 3 components
```{r}
bzsg_items <- ee_final %>% select(bzsg1, bzsg2, bzsg3)

mean(bzsg_items$bzsg1)
sd(bzsg_items$bzsg1)

mean(bzsg_items$bzsg2)
sd(bzsg_items$bzsg2)

mean(bzsg_items$bzsg3)
sd(bzsg_items$bzsg3)
```

###Mean and SD of Theft components
```{r}
theft_items <- ee_final %>% select(theft1, theft2, theft3)

mean(theft_items$theft1)
sd(theft_items$theft1)

mean(theft_items$theft2)
sd(theft_items$theft2)

mean(theft_items$theft3)
sd(theft_items$theft3)
```
###Mean and SD of BJW 5 components
```{r}
bjw_items <- ee_final %>% select(bjw_nature, bjw_god, bjw_nr, bjw_deserve, bjw_chance)

mean(bjw_items$bjw_nature)
sd(bjw_items$bjw_nature)

mean(bjw_items$bjw_god)
sd(bjw_items$bjw_god)

mean(bjw_items$bjw_nr)
sd(bjw_items$bjw_nr)

mean(bjw_items$bjw_deserve)
sd(bjw_items$bjw_deserve)

mean(bjw_items$bjw_chance)
sd(bjw_items$bjw_chance)
```
#Additional analyses
###Cronbach's Alpha
```{r}
#interpreting cronbach's alpha
#http://pioneer.netserv.chula.ac.th/~ppongsa/2013605/Cronbach.pdf

#As a rule of thumb, a Cronbach’s alpha of .50 – .80 is usually considered an acceptable value (Field, 2010; Setbon & Raude, 2010), with considerably lower values indicating lack of reliability.

#http://www.neilmclatchie.com/using-r-cronbachs-alpha/
#https://www.bing.com/videos/search?q=making+table+for+cronbach+alpha+results&&view=detail&mid=4E5B62E3AED97E71CD0A4E5B62E3AED97E71CD0A&&FORM=VRDGAR&ru=%2Fvideos%2Fsearch%3Fq%3Dmaking%2Btable%2Bfor%2Bcronbach%2Balpha%2Bresults%26qs%3Dn%26form%3DQBVR%26sp%3D-1%26pq%3Dmaking%2Btable%2Bfor%2Bcronbach%2Balpha%2Bresult%26sc%3D0-38%26sk%3D%26cvid%3D1E433D4AEB6E40558DEBA38161BED665


ee_i_m <- data.matrix(ee_items)

is.matrix(ee_i_m)
alpha(ee_items, n.var= 10)
?alpha

install.packages("psych")
library(psych)

#https://www.bing.com/videos/search?q=cronbach%27s+alpha+in+r+youtube&docid=608024166374967704&mid=4E9425AC8986FA0195664E9425AC8986FA019566&view=detail&FORM=VIRE

#2 methods

install.packages("umx")
library(umx)

#figuring it out with EE

ee_i_m <- data.matrix(ee_items)

is.matrix(ee_i_m)

reliability(cov(ee_i_m))

?reliability
#above did not work?

#2nd method
#alpha using psych 
library(psych)
#alpha(ee_i_m)
#the output: what the actual fuck


#BELOW WORKS
ee_i_m <- data.matrix(ee_items)

is.data.frame(ee_items)

is.matrix(ee_i_m)

umx::reliability(cov(ee_i_m))

#EE


#CII

umx::reliability(cov(cii_items))
#SRD
umx::reliability(cov(srd_items))
#Theft
umx::reliability(cov(theft_items))

#BJW
umx::reliability(cov(bjw_items))

#BZSG
umx::reliability(cov(bzsg_items))

#II

```
##Correlation Matrix
```{r}
#new correlation matrix 
cm_lm1_r <- ee_final %>% dplyr::select(c(ee_sum, bzsg_sum, cii, theft_sum, bjw_sum, srd_sum, ii_sum))

vars <- data.matrix(cm_lm1_r)
var_t <- cor(vars)

View(var_t)

class(var_t)
```
##New Correlation Matrix graphs
```{r}

#colorblind friendly graphs from package
#

https://www.rdocumentation.org/packages/colorblindr


#info on plots
#http://jamesmarquezportfolio.com/correlation_matrices_in_r.html

#GGsally

var_t <- as.data.frame(var_t)
viz_var_t <- ggpairs(var_t)
View(var_t)
?ggpairs
var_num <- var_t[,]
viz_var_t
View(viz_var_t)

#example from site
#ggcorr(mydata, nbreaks=8, palette='RdGy', label=TRUE, label_size=5, label_color='white')

ggcorr(var_t, nbreaks=8, palette='RdGy', label=TRUE, label_size=2, label_color='white')


#another example from site
library(ggcorrplot)

ggcorrplot(cor(var_t), p.mat = cor_pmat(var_t), hc.order=TRUE, type='lower')

ggcorrplot(cor(var_t), hc.order=TRUE, type='lower')


#correlations
ee_bjw <- ee_final[,24:28]
ee_bjw$ee_sum <- ee_final$ee_sum
View(ee_bjw)

#ee and bjw components

ggcorr(ee_bjw, nbreaks=8, palette='RdGy', label=TRUE, label_size=2, label_color='white')


#just using key numeric variables



cm_lm1_r <- ee_final %>% dplyr::select(c(ee_sum, bzsg_sum, cii, theft_sum, bjw_sum, srd_sum, ii_sum))

vars <- data.matrix(cm_lm1_r)



ggcorr(vars, nbreaks=8, palette='RdGy', label=TRUE, label_size=2, label_color='white')

View(ee_final)

#bzsg and ee

bzsg_ee <- ee_final %>% select(bzsg1, bzsg2, bzsg3, ee_sum)

ggcorr(bzsg_ee, nbreaks=8, palette='RdGy', label=TRUE, label_size=2, label_color='white')

#srd
srd_ee <- ee_final %>% select(srd1, srd2, srd3_st, srd4_st, ee_sum)

ggcorr(srd_ee, nbreaks=8, palette='RdGy', label=TRUE, label_size=2, label_color='white')

#I feel deprived was best predictor of EE_sum from srd (2)

#ii

ii_ee <- ee_final %>% select(ii1_st, ii2, ee_sum)

ggcorr(ii_ee, nbreaks=8, palette='RdGy', label=TRUE, label_size=2, label_color='white')

#awkwardly my two items are not super correlated (.3)

#cii
cii_ee <- ee_final %>% select(police1, police2, courts1, courts2, ee_sum)

ggcorr(cii_ee, nbreaks=8, palette='RdGy', label=TRUE, label_size=2, label_color='white')

#theft
theft_ee <- ee_final %>% select(theft1, theft2, theft3, ee_sum)

ggcorr(theft_ee, nbreaks=8, palette='RdGy', label=TRUE, label_size=2, label_color='white')
View(ee_final)

View(ee_final)

#correlations between summaries of elements within the final model

f_m <- ee_final %>% select(ee_sum, children, srd_sum, theft_sum, cii, bjw_sum)
ggcorr(f_m, nbreaks=8, palette='RdGy', label=TRUE, label_size=2, label_color='white')
View(ee_final)
```




